---
title: "02-SummaryStats"
author: "Yifei Ding"
date: "2024-04-09"
output: html_document
---
```{r}
Sys.setenv(lang = "en")
```

# Working With Data  {#sum}


In this chapter we will first learn some basic concepts that help summarizing data. Then, we will tackle a real-world task and read, clean, and summarize data from the web.

## Summary Statistics

`R` has built in functions for a large number of summary statistics. For numeric variables, we can summarize data by looking at their center and spread, for example.

```{r}
# for the mpg dataset, we load:
library(ggplot2)
```

```{r}
head(mpg)
```

### Central Tendency {-}

Suppose we want to know the *mean* and *median* of all the values stored in the `data.frame` column `mpg$cty`:

| Measure | `R`               | Result              |
|:---------:|:-------------------:|:---------------------:|
| Mean    | `mean(mpg$cty)`   | `r mean(mpg$cty)`   |
| Median  | `median(mpg$cty)` | `r median(mpg$cty)` |

### Spread {-}

How do the values in that column *vary*? How far *spread out* are they?

| Measure            | `R`              | Result             |
|:---------:|:-------------------:|:---------------------:|
| Variance           | `var(mpg$cty)`   | `r var(mpg$cty)`   |
| Standard Deviation | `sd(mpg$cty)`    | `r sd(mpg$cty)`    |
| IQR                | `IQR(mpg$cty)`   | `r IQR(mpg$cty)`   |
| Minimum            | `min(mpg$cty)`   | `r min(mpg$cty)`   |
| Maximum            | `max(mpg$cty)`   | `r max(mpg$cty)`   |
| Range              | `range(mpg$cty)` | `r range(mpg$cty)` |

### Categorical {-}

For categorical variables, counts and percentages can be used for summary.

```{r}
table(mpg$drv)
table(mpg$drv) / nrow(mpg)
```
